---
title: "HolyDead Feedback Analysis"
subtitle: "Report from production phase of user testing"
author: "Dawood Parhiar"
date: "04/22/2025"
output: pdf
editor: visual
---

```{r setup, include=FALSE}
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("readxl")) install.packages("readxl")
if (!require("janitor")) install.packages("janitor")
if (!require("wordcloud")) install.packages("wordcloud")
if (!require("tm")) install.packages("tm")

library(tidyverse)
library(readxl)
library(janitor)
library(wordcloud)
library(tm)

survey_data <- read_excel("Data/Holy_Dead_Feedback_Production.xlsx")
survey_data <- survey_data %>% clean_names()


```

```{r setup , include=FALSE}
names(survey_data)
```

# Introduction

This report presents a descriptive analysis of data collected from user testing of our game Holy Dead in the production phase. Participants answered questions about the appeal, controls and fun factor of the game.

We use basic statistics and visualizations to explore relationships between variables and identify patterns.

------------------------------------------------------------------------

# 1. Dataset Overview

We begin by exploring the structure of the dataset.

## 1.1 Preview of the Data

```{r , include=FALSE}
head(survey_data)
```

## 1.2 Dataset Structure

```{r , include=FALSE}
str(survey_data)
```

#2. Key variables descriptive statistics summary Here we summarize numerical variables using measures like mean, median, and standard deviation.

```{r}
survey_data %>%
  mutate(
    Age_group = as.factor(what_is_your_age_group),
    Gameplay_Time = as.factor(how_long_did_you_play_the_game),
    Difficulty = as.factor(was_the_game_difficulty_balanced)
  ) %>%
  select(Age_group, Gameplay_Time, on_a_scale_of_1_5_how_enjoyable_was_the_game, Difficulty, how_responsive_did_the_controls_feel_1_laggy_5_very_responsive) %>%
  summary()

```

# 3.1 Enjoyability & Game Duration

```{r}

# boxplot with outlier detection for play time categories
ggplot(survey_data, aes(x = how_long_did_you_play_the_game, 
                        y = as.numeric(on_a_scale_of_1_5_how_enjoyable_was_the_game))) +
  geom_boxplot(fill = "darkcyan", outlier.color = "red", outlier.shape = 8) +
  labs(title = "Enjoyability by Play Time (Boxplot)", 
       x = "Play Time Duration", 
       y = "Enjoyment Rating (1–5)") +
  theme_minimal()


```

# 2.1 Controls and UI

```{r}
ggplot(survey_data, aes(x = were_the_controls_easy_to_learn)) +
  geom_bar(fill = "lightblue") +
  labs(title = "Ease of Learning Controls", x = "Response", y = "Count")

survey_data %>%
  ggplot(aes(x = how_responsive_did_the_controls_feel_1_laggy_5_very_responsive)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "white") +
  labs(title = "Control Responsiveness", x = "Rating (1 = Laggy, 5 = Very Responsive)", y = "Count")

ggplot(survey_data, aes(x = did_the_games_ui_elements_feel_clear_and_intuitive)) +
  geom_bar(fill = "steelblue") +
  labs(title = "Clarity of UI Elements", x = "Response", y = "Count")
```

# 3. Challenge and Confusion

```{r}
ggplot(survey_data, aes(x = was_the_game_difficulty_balanced)) +
  geom_bar(fill = "orange") +
  labs(title = "Was the Game Difficulty Balanced?", x = "Response", y = "Count")

ggplot(survey_data, aes(x = did_you_get_stuck_or_confused_at_any_point)) +
  geom_bar(fill = "tomato") +
  labs(title = "Did Players Get Stuck or Confused?", x = "Response", y = "Count")
```

# 4. Immersion and Experience

```{r}
ggplot(survey_data, aes(x = did_the_game_audio_enhance_your_experience)) +
  geom_bar(fill = "darkgreen") +
  labs(title = "Did Audio Enhance the Experience?", x = "Response", y = "Count")

ggplot(survey_data, aes(x = did_the_game_world_feel_immersive)) +
  geom_bar(fill = "forestgreen") +
  labs(title = "Immersion in Game World", x = "Response", y = "Count")

ggplot(survey_data, aes(x = did_you_feel_a_sense_of_progression_as_you_played)) +
  geom_bar(fill = "seagreen") +
  labs(title = "Sense of Progression", x = "Response", y = "Count")
```

# 5. Visual & Audio Fidelity

```{r}
ggplot(survey_data, aes(x = did_the_character_movement_feel_smooth_and_natural)) +
  geom_bar(fill = "purple") +
  labs(title = "Character Movement Feedback", x = "Response", y = "Count")

ggplot(survey_data, aes(x = did_the_graphics_and_art_style_match_the_games_theme)) +
  geom_bar(fill = "darkred") +
  labs(title = "Visual Style Fit the Theme?", x = "Response", y = "Count")
```

# 6. Bugs, Glitches & Performance

```{r}
ggplot(survey_data, aes(x = did_you_encounter_any_bugs_or_glitches)) +
  geom_bar(fill = "red") +
  labs(title = "Bug or Glitch Encountered?", x = "Response", y = "Count")

ggplot(survey_data, aes(x = did_you_experience_any_lag_frame_drops_or_performance_issues)) +
  geom_bar(fill = "firebrick") +
  labs(title = "Performance Issues Experienced?", x = "Response", y = "Count")
```

# 7. Word Clouds (Open Feedback)

```{r}
least_enjoyed_text <- na.omit(survey_data$what_part_of_the_game_did_you_enjoy_the_least)
if (length(least_enjoyed_text) > 0) {
  corpus <- Corpus(VectorSource(least_enjoyed_text)) %>%
    tm_map(content_transformer(tolower)) %>%
    tm_map(removePunctuation) %>%
    tm_map(removeNumbers) %>%
    tm_map(removeWords, stopwords("en"))

  wordcloud(corpus, max.words = 100, colors = brewer.pal(8, "Dark2"), random.order = FALSE)
} else {
  print("No responses for least enjoyed game part.")
}

bugs_text <- na.omit(survey_data$if_yes_please_specify_2)
if (length(bugs_text) > 0) {
  corpus2 <- Corpus(VectorSource(bugs_text)) %>%
    tm_map(content_transformer(tolower)) %>%
    tm_map(removePunctuation) %>%
    tm_map(removeNumbers) %>%
    tm_map(removeWords, stopwords("en"))

  wordcloud(corpus2, max.words = 100, colors = brewer.pal(8, "Dark2"), random.order = FALSE)
} else {
  print("No bug description responses.")
}
```

# 8. Boxplots & Correlation Insights

## 8.1 Boxplot: Enjoyability vs. Play Time

```{r}

# boxplot: Enjoyment rating by play time
ggplot(survey_data, aes(x = how_long_did_you_play_the_game,
                        y = as.numeric(on_a_scale_of_1_5_how_enjoyable_was_the_game))) +
  geom_boxplot(fill = "plum", alpha = 0.7) +
  labs(title = "Enjoyability by Play Time",
       x = "Play Time",
       y = "Enjoyment Rating (1–5)") +
  theme_minimal()


 
```

```{r}
ggplot(survey_data, aes(x = as.factor(how_responsive_did_the_controls_feel_1_laggy_5_very_responsive),
                        y = as.numeric(on_a_scale_of_1_5_how_enjoyable_was_the_game))) +
  geom_boxplot(fill = "skyblue", alpha = 0.7) +
  labs(title = "Enjoyability vs. Control Responsiveness",
       x = "Control Responsiveness (1–5)",
       y = "Enjoyment Rating (1–5)") +
  theme_minimal()
```

# Conclusion

This visual report captures the full spectrum of feedback from the HolyDead survey, helping identify strong and weak areas in gameplay, design, and overall experience.
