---
title: "HolyDead Feedback Analysis"
subtitle: "Full Visual Report of Player Responses"
author: "Dawood Parhiar"
date: "04/22/2025"
output: html_document
---

```{r setup, include=FALSE}
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("readxl")) install.packages("readxl")
if (!require("janitor")) install.packages("janitor")
if (!require("wordcloud")) install.packages("wordcloud")
if (!require("tm")) install.packages("tm")

library(tidyverse)
library(readxl)
library(janitor)
library(wordcloud)
library(tm)

survey_data <- read_excel("Data/Feedback1.xlsx")
survey_data <- survey_data %>% clean_names()


```

```{r}
names(survey_data)
```
# Introduction

This report presents a descriptive analysis of data collected from a survey on streaming habits. Participants answered questions about their viewing behavior, mood, and platform preferences.

We use basic statistics and visualizations to explore relationships between variables and identify patterns.

---

# 1. Dataset Overview

We begin by exploring the structure of the dataset.

## 1.1 Preview of the Data

```{r}
head(survey_data)
```

## 1.2 Dataset Structure

```{r}
str(survey_data)
```


# 1. Enjoyability & Game Duration

```{r}
ggplot(survey_data, aes(x = as.factor(on_a_scale_of_1_5_how_enjoyable_was_the_game))) +
  geom_bar(fill = "darkorchid") +
  labs(title = "Game Enjoyability (1–5 Scale)", x = "Enjoyment Rating", y = "Count")

# Count ratings
enjoy_counts <- survey_data %>%
  count(on_a_scale_of_1_5_how_enjoyable_was_the_game) %>%
  mutate(percentage = n / sum(n),
         label = paste0(on_a_scale_of_1_5_how_enjoyable_was_the_game, " (", round(percentage * 100), "%)"))

# donut chart
ggplot(enjoy_counts, aes(x = 2, y = percentage, fill = as.factor(on_a_scale_of_1_5_how_enjoyable_was_the_game))) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  xlim(0.5, 2.5) +  # Creates the donut hole
  theme_void() +
  geom_text(aes(label = label), position = position_stack(vjust = 0.5), size = 4) +
  scale_fill_brewer(palette = "Dark2") +
  labs(title = "Game Enjoyability (Donut Chart)", fill = "Rating")


survey_data %>%
  count(how_long_did_you_play_the_game) %>%
  ggplot(aes(x = reorder(how_long_did_you_play_the_game, -n), y = n)) +
  geom_col(fill = "darkcyan") +
  labs(title = "How Long Did Players Play the Game?", x = "Play Time", y = "Count")

# boxplot with outlier detection for play time categories
ggplot(survey_data, aes(x = how_long_did_you_play_the_game, 
                        y = as.numeric(on_a_scale_of_1_5_how_enjoyable_was_the_game))) +
  geom_boxplot(fill = "darkcyan", outlier.color = "red", outlier.shape = 8) +
  labs(title = "Enjoyability by Play Time (Boxplot)", 
       x = "Play Time Duration", 
       y = "Enjoyment Rating (1–5)") +
  theme_minimal()


```

# 2. Controls and UI

```{r}
ggplot(survey_data, aes(x = were_the_controls_easy_to_learn)) +
  geom_bar(fill = "lightblue") +
  labs(title = "Ease of Learning Controls", x = "Response", y = "Count")

survey_data %>%
  ggplot(aes(x = how_responsive_did_the_controls_feel_1_laggy_5_very_responsive)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "white") +
  labs(title = "Control Responsiveness", x = "Rating (1 = Laggy, 5 = Very Responsive)", y = "Count")

ggplot(survey_data, aes(x = did_the_games_ui_elements_feel_clear_and_intuitive)) +
  geom_bar(fill = "steelblue") +
  labs(title = "Clarity of UI Elements", x = "Response", y = "Count")
```

# 3. Challenge and Confusion

```{r}
ggplot(survey_data, aes(x = was_the_game_difficulty_balanced)) +
  geom_bar(fill = "orange") +
  labs(title = "Was the Game Difficulty Balanced?", x = "Response", y = "Count")

ggplot(survey_data, aes(x = did_you_get_stuck_or_confused_at_any_point)) +
  geom_bar(fill = "tomato") +
  labs(title = "Did Players Get Stuck or Confused?", x = "Response", y = "Count")
```

# 4. Immersion and Experience

```{r}
ggplot(survey_data, aes(x = did_the_game_audio_enhance_your_experience)) +
  geom_bar(fill = "darkgreen") +
  labs(title = "Did Audio Enhance the Experience?", x = "Response", y = "Count")

ggplot(survey_data, aes(x = did_the_game_world_feel_immersive)) +
  geom_bar(fill = "forestgreen") +
  labs(title = "Immersion in Game World", x = "Response", y = "Count")

ggplot(survey_data, aes(x = did_you_feel_a_sense_of_progression_as_you_played)) +
  geom_bar(fill = "seagreen") +
  labs(title = "Sense of Progression", x = "Response", y = "Count")
```

# 5. Visual & Audio Fidelity

```{r}
ggplot(survey_data, aes(x = did_the_character_movement_feel_smooth_and_natural)) +
  geom_bar(fill = "purple") +
  labs(title = "Character Movement Feedback", x = "Response", y = "Count")

ggplot(survey_data, aes(x = did_the_graphics_and_art_style_match_the_games_theme)) +
  geom_bar(fill = "darkred") +
  labs(title = "Visual Style Fit the Theme?", x = "Response", y = "Count")
```

# 6. Bugs, Glitches & Performance

```{r}
ggplot(survey_data, aes(x = did_you_encounter_any_bugs_or_glitches)) +
  geom_bar(fill = "red") +
  labs(title = "Bug or Glitch Encountered?", x = "Response", y = "Count")

ggplot(survey_data, aes(x = did_you_experience_any_lag_frame_drops_or_performance_issues)) +
  geom_bar(fill = "firebrick") +
  labs(title = "Performance Issues Experienced?", x = "Response", y = "Count")
```

# 7. Word Clouds (Open Feedback)

```{r}
least_enjoyed_text <- na.omit(survey_data$what_part_of_the_game_did_you_enjoy_the_least)
if (length(least_enjoyed_text) > 0) {
  corpus <- Corpus(VectorSource(least_enjoyed_text)) %>%
    tm_map(content_transformer(tolower)) %>%
    tm_map(removePunctuation) %>%
    tm_map(removeNumbers) %>%
    tm_map(removeWords, stopwords("en"))

  wordcloud(corpus, max.words = 100, colors = brewer.pal(8, "Dark2"), random.order = FALSE)
} else {
  print("No responses for least enjoyed game part.")
}

bugs_text <- na.omit(survey_data$if_yes_please_specify_2)
if (length(bugs_text) > 0) {
  corpus2 <- Corpus(VectorSource(bugs_text)) %>%
    tm_map(content_transformer(tolower)) %>%
    tm_map(removePunctuation) %>%
    tm_map(removeNumbers) %>%
    tm_map(removeWords, stopwords("en"))

  wordcloud(corpus2, max.words = 100, colors = brewer.pal(8, "Dark2"), random.order = FALSE)
} else {
  print("No bug description responses.")
}
```
# 8. Boxplots & Correlation Insights

## 8.1 Boxplot: Enjoyability vs. Play Time

```{r}

# boxplot: Enjoyment rating by play time
ggplot(survey_data, aes(x = how_long_did_you_play_the_game,
                        y = as.numeric(on_a_scale_of_1_5_how_enjoyable_was_the_game))) +
  geom_boxplot(fill = "plum", alpha = 0.7) +
  labs(title = "Enjoyability by Play Time",
       x = "Play Time",
       y = "Enjoyment Rating (1–5)") +
  theme_minimal()


 
```

```{r}
ggplot(survey_data, aes(x = as.factor(how_responsive_did_the_controls_feel_1_laggy_5_very_responsive),
                        y = as.numeric(on_a_scale_of_1_5_how_enjoyable_was_the_game))) +
  geom_boxplot(fill = "skyblue", alpha = 0.7) +
  labs(title = "Enjoyability vs. Control Responsiveness",
       x = "Control Responsiveness (1–5)",
       y = "Enjoyment Rating (1–5)") +
  theme_minimal()
```

# Conclusion

This visual report captures the full spectrum of feedback from the HolyDead survey, helping identify strong and weak areas in gameplay, design, and overall experience.
