---
title: "HolyDead Feedback Analysis"
subtitle: "Report 2 from production phase of user testing "
author: "Dawood Parhiar"
date: "23/04/2025"
format: pdf 
editor: visual
---

```{r setup, include=FALSE}
# Load libraries
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("readxl")) install.packages("readxl")
if (!require("janitor")) install.packages("janitor")
if (!require("wordcloud")) install.packages("wordcloud")
if (!require("tm")) install.packages("tm")

library(tidyverse)
library(readxl)
library(janitor)
library(wordcloud)
library(tm)

# Load and clean data
survey_data <- read_excel("Data/Feedback2.xlsx") %>%
  clean_names()
```

## 1. Dataset Overview

```{r , include=FALSE}
head(survey_data)
str(survey_data)
```

## 2. Fun Factor Rating (Donut Chart)

```{r}
fun_counts <- survey_data %>%
  count(how_well_would_you_rate_your_overall_fun_factor_of_playing_this_game) %>%
  mutate(percentage = n / sum(n),
         label = paste0(how_well_would_you_rate_your_overall_fun_factor_of_playing_this_game, " (", round(percentage*100), "%)"))

ggplot(fun_counts, aes(x = 2, y = percentage, fill = as.factor(how_well_would_you_rate_your_overall_fun_factor_of_playing_this_game))) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  xlim(0.5, 2.5) +
  theme_void() +
  geom_text(aes(label = label), position = position_stack(vjust = 0.5)) +
  scale_fill_brewer(palette = "Dark2") +
  labs(title = "Fun Factor Rating (1–5)", fill = "Rating")
```

## 3. Fun Factor by Age Group (Boxplot)

```{r}
ggplot(survey_data, aes(x = what_is_your_age_group , y = as.numeric(how_well_would_you_rate_your_overall_fun_factor_of_playing_this_game))) +
  geom_boxplot(fill = "lightblue", outlier.color = "red", outlier.shape = 8) +
  labs(title = "Fun Factor Rating by Age Group", x = "Age Group", y = "Rating (1–5)") +
  theme_minimal()
```

## 4. Word Cloud: What Did Players Like Most?

```{r}
liked_text <- na.omit(survey_data$what_aspects_of_the_game_did_you_like_most)
if (length(liked_text) > 0) {
  corpus <- Corpus(VectorSource(liked_text)) %>%
    tm_map(content_transformer(tolower)) %>%
    tm_map(removePunctuation) %>%
    tm_map(removeNumbers) %>%
    tm_map(removeWords, stopwords("en"))
  
  wordcloud(corpus, max.words = 100, colors = brewer.pal(8, "Dark2"))
} else {
  print("No responses available for this section.")
}
```

## 5. Word Cloud: Suggestions for Improvement

```{r}
improve_text <- na.omit(survey_data$what_would_make_this_game_more_exciting_in_your_opinion)
if (length(improve_text) > 0) {
  corpus2 <- Corpus(VectorSource(improve_text)) %>%
    tm_map(content_transformer(tolower)) %>%
    tm_map(removePunctuation) %>%
    tm_map(removeNumbers) %>%
    tm_map(removeWords, stopwords("en"))

  wordcloud(corpus2, max.words = 100, colors = brewer.pal(8, "Dark2"))
} else {
  print("No suggestions recorded.")
}
```

## 6. Bug Reports Summary

```{r}
bug_text <- na.omit(survey_data$if_you_encountered_any_bugs_in_the_game_play_please_state_below)
if (length(bug_text) > 0) {
  corpus3 <- Corpus(VectorSource(bug_text)) %>%
    tm_map(content_transformer(tolower)) %>%
    tm_map(removePunctuation) %>%
    tm_map(removeNumbers) %>%
    tm_map(removeWords, stopwords("en"))

  wordcloud(corpus3, max.words = 100, colors = brewer.pal(8, "Dark2"), random.order = FALSE)
} else {
  print("No bug reports submitted.")
}
```

## Conclusion

This analysis summarizes key sentiments and suggestions from players based on the second feedback round for HolyDead. The visualizations provide insights into fun factor perception, frequent issues, and improvement ideas to iterate on the next version.
