---
title: "Survey Analysis Holy Dead"
subtitle: "Post Production User Testing"
Author: "Dawood Parhiar"
date: "22/04/2025"
---

```{r setup, include=FALSE}
# Load tidyverse for wrangling and visualization

if (!require("tidyverse")) install.packages("tidyverse")
if (!require("readxl")) install.packages("readxl")
if (!require("janitor")) install.packages("janitor")
if (!require("wordcloud")) install.packages("wordcloud")
if (!require("tm")) install.packages("tm")

library(tidyverse)
library(readxl)
library(janitor)
library(wordcloud)
library(tm)

# For data wrangling, plotting — includes ggplot2, dplyr, readr, etc.

# Clear Console
cat("\014")

library(wordcloud)

# Load survey data

survey_data <- read_excel("Data/Holy_Dead_Feedback_Form_Post_Production.xlsx") 

survey_data <- survey_data %>% clean_names()
```
#Introduction
This report presents a descriptive analysis of data collected from a user testing session that was conducted with the mixed audience who played our game HolyDead and answered questions through a feedback form.

We use basic statistics and visualizations to explore relationships between variables and identify patterns.

---

# 1. Dataset Overview

We begin by exploring the structure of the dataset.

## 1.1 Preview of the Data

```{r}
head(survey_data)
```

## 1.2 Dataset Structure

```{r}
str(survey_data)
```

```{r}
names(survey_data)
```
---


# 2. Descriptive Statistics

## 2.1 Enjoyability Summary

```{r}

table(survey_data$on_a_scale_of_1_5_how_enjoyable_was_the_game)
```

## 2.2 Control Responsiveness

```{r}
summary(survey_data$how_responsive_did_the_controls_feel_1_laggy_5_very_responsive)
```


## 3.2 Control Responsiveness (Histogram)

```{r}
ggplot(survey_data, aes(x = how_responsive_did_the_controls_feel_1_laggy_5_very_responsive)) +
  geom_histogram(binwidth = 1, fill = "purple", color = "white") +
  labs(title = "Control Responsiveness",
       x = "Responsiveness Rating",
       y = "Number of Responses")
```

## 3.3 Immersion Feedback (Pie Chart)

```{r}
library(ggplot2)
library(dplyr)

data <- survey_data %>%
  count(did_the_game_world_feel_immersive) %>%
  mutate(prop = n / sum(n),
         ypos = cumsum(prop) - 0.5 * prop)

ggplot(data, aes(x = 2, y = prop, fill = did_the_game_world_feel_immersive)) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  xlim(0.5, 2.5) +
  geom_text(aes(y = ypos, label = scales::percent(prop, accuracy = 1)), color = "white") +
  theme_void() +
  labs(title = "Immersion Feedback (Donut Chart)") +
  theme(legend.position = "right")


```

## 3.4 Enjoyability by Age Group (Boxplot)

```{r}
ggplot(survey_data, aes(x = what_is_our_age_group,
                        y = as.numeric(on_a_scale_of_1_5_how_enjoyable_was_the_game))) +
  geom_boxplot(fill = "orange") +
  labs(title = "Enjoyability by Age Group",
       x = "Age Group",
       y = "Enjoyment Rating")
```

---

# 4. Bugs and Glitches Feedback

## 4.1 Bug Reports (Bar Chart)

```{r}
ggplot(survey_data, aes(x = did_you_encounter_any_bugs_or_glitches)) +
  geom_bar(fill = "red") +
  labs(title = "Reported Bugs or Glitches",
       x = "Response",
       y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

---
# 5. Word Clouds for Open Feedback

## 5.1 Bug Descriptions

```{r}
bugs_text <- survey_data$if_yes_please_specify
bugs_corpus <- Corpus(VectorSource(bugs_text))
bugs_corpus <- bugs_corpus %>%
  tm_map(content_transformer(tolower)) %>%
  tm_map(removePunctuation) %>%
  tm_map(removeNumbers) %>%
  tm_map(removeWords, stopwords("en"))

wordcloud(bugs_corpus, max.words = 100, colors = brewer.pal(8, "Dark2"), random.order = FALSE)
```

## 5.2 Least Enjoyed Game Parts

```{r}
least_enjoyed_text <- survey_data$what_part_of_the_game_did_you_enjoy_the_least
least_enjoyed_corpus <- Corpus(VectorSource(least_enjoyed_text)) %>%
  tm_map(content_transformer(tolower)) %>%
  tm_map(removePunctuation) %>%
  tm_map(removeNumbers) %>%
  tm_map(removeWords, stopwords("en"))

wordcloud(least_enjoyed_corpus, max.words = 100, colors = brewer.pal(8, "Dark2"), random.order = FALSE)
```

# Conclusion

This analysis highlights general player sentiment and usability of the HolyDead game. Most players found the controls responsive and the game immersive, with a few areas such as clarity of feedback and minor bugs noted for improvement.
